### 2021-12-06 submission/manuscript.rmd
add risk stratified results and update methods

---
title: "Treatment heterogeneity in comparative effectiveness of teriparatide vs bisphosphonates: multi-database cohort study"
output:
    bookdown::pdf_document2: default
    bookdown::word_document2:
        reference_docx: reference.docx
geometry: margin=1.0in
toc: false
font-size: 11pt
header-includes:
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage[left]{lineno}
editor_options: 
  chunk_output_type: console
bibliography: references.bib
csl: jamia.csl
---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(here)

options(scipen=1, digits=2)

d <- function(x, decimals = 2) {
  sprintf(paste0("%1.", decimals, "f"), x) 
}

knit_hooks$set(
  inline = function(x) {
    prettyNum(
      x,
      big.mark = ",",
      decimal.mark = ".",
      preserve.width = "individual"
    )
  }
)
```

\vspace{35mm}

Alexandros Rekkas${^1}$, Annika M. J&ouml;dicke${^2}$, David van Klaveren$^3$,
Daniel Prieto-Alhambra${^{1,2}}$, Peter R. Rijnbeek${^1}$

\vspace{40mm}

$^1$ Department of Medical Informatics, Erasmus University Medical Center,
Rotterdam, Netherlands

$^2$ Pharmaco- and Device Epidemiology, Centre for Statistics in Medicine,
Nuffield Department of Orthopaedics, Rheumatology and Musculoskeletal Sciences,
University of Oxford, Oxford, UK

$^3$ Department of Public Health, Erasmus University Medical Center, Rotterdam,
Netherlands

\newpage

\linenumbers
# Introduction
Osteoporosis is a common condition characterised by decreased bone
density and associated with increased risk for fragility fractures,
which affects almost 30% of women aged $\geq50$ years
[@wright2017impact]. It was estimated that in 2005, over 2 million
incident fractures occurred in the United States, with hip fractures
accounting for more than 70% of the costs. By 2025, annual fractures
and costs could rise by almost 50% [@Cole2008].

Fracture prevention is thus the key focus of anti-osteoporotic
therapy, reducing the disease burden both for the affected patient but
also on the population level. Several pharmacological agents are
available for primary and secondary prevention. The choice of
anti-osteoporotic agent largely depends on history of fragility
fracture and anticipated fracture risk. Oral bisphosphonates (BP) are
first line treatments for postmenopausal patients with increased
fracture risk in clinical guidelines (#AACE/ACE 2020) as for their
favourable cost-effectiveness and safety profile. Teriparatide, a
parathyroid hormone analogue administered as a daily injection, was
approved by the FDA in 2002 as the first anabolic agent for treatment
of severe postmenopausal osteoporosis. While clinical trials among
patients with previous vertebral fractures showed a substantial risk
reduction for new vertebral fracture comparing teriparatide to placebo
[@Neer2001] and risedronate [@Kendler2018], its efficacy on
low-incident major osteoporotic fractures, especially hip fractures,
is less well established: Previous randomised trials assessing hip
fracture comprised only few events, thus not providing sufficient
power to show differences between treatment groups when studying
comparative effectiveness. A recent meta-analysis by Diez-Perez and
colleagues assessed the effect of teriparatide on hip fracture,
indicating a significant 80% risk reduction compared to placebo and a
non-significant 46% risk reduction when compared to active controls
[@DezPrez2019].

With teriparatide-containing biosimilars being launched in recent
years, treatment costs dropped significantly. Subsequently, the
discussion of cost-effectiveness in patients with less severe forms of
osteoporosis is restarting. Therefore, evaluation of the effectiveness
in hip fracture prevention is required. In particular, observational
studies assessing comparative effectiveness in a real-world setting
are required, as effectiveness may differ from clinical trial as for
differences in patient’s persistence and adherence to their
anti-osteoporotic treatment. While hip fracture is a comparatively
rare outcome, it is unambiguously defined and reliably recorded in
routinely collected data, allowing for a direct comparison to results
from clinical trial meta-analyses. While results from non-controlled,
observational studies suggest that teriparatide may reduce the risk of
hip fractures [@Burge2016; @Silverman2018], this is the first
observational study assessing teriparatide in hip fracture risk
reduction compared to oral bisphosphonate users in postmenopausal
women using multiple large real-world databases.

\newpage
# Methods
## Study design
We performed a retrospective new user cohort study to estimate the
effectiveness of teriparatide compared to oral bisphosphonates in
patients with osteoporosis[8]. We included female participants above
the age of 50 with established osteoporosis (any condition occurrence
of hip, wrist, spine or shoulder/humerus fracture in their history,
prior to treatment initiation) and at least 365 days of continuous
observation period before the index event. Patients were considered to
be new-users if they received no anti-osteoporosis drugs (raloxifene,
bazedoxifene, denosumab, abaloparatide, romosozumab) 365 days prior to
treatment initiation with teriparatide or an oral bisphosphonate.

Our primary efficacy outcome was hip fracture. Vertebral fracture and
a composite major osteoporotic fracture, defined as hip, vertebral or
wrist/forearm/proximal humerus fracture, were our secondary efficacy
outcomes. We excluded patients that had experienced the outcome under
study any time prior to treatment initiation. Patient time-at-risk
started 1 day after treatment initiation and finished 730 days after
treatment initiation.

## Data sources

We ran our analyses on three US observational databases mapped to OMOP-CDM
version. More specifically:

- **IMB MarketScan Commercial Database (CCAE)** includes health
  insurance claims across the continuum of care (e.g. inpatient,
  outpatient, outpatient pharmacy, carve-out behavioral healthcare) as
  well as enrollment data from large employers and health plans across
  the United States who provide private healthcare coverage for more
  than 155 million employees, their spouses, and dependents. This
  administrative claims database includes a variety of fee-
  for-service, preferred provider organizations, and capitated health
  plans.
- **Optum De-Identified Clinformatics® Data Mart Database – Date of
  Death (Optum-DOD)**: is derived from a database of administrative
  health claims for members of large commercial and Medicare Advantage
  health plans. The database includes approximately 17-19 million
  annual covered lives, for a total of over 65 million unique lives
  over a 12 year period (1/2007 through 12/2019). The population is
  geographically diverse, spanning all 50 states.
- **Optum de-identified Electronic Health Record Dataset (Optum-EHR)**
  is derived from dozens of healthcare provider organizations in the
  United States, that include more than 700 Hospitals and 7000
  Clinics; treating more than 102 million patients receiving care in
  the United States.

##     Statistical analyses
We carried out two sets of analyses. First, we derived overall
treatment effect estimates of teriparatide compared to oral
bisphosphonates regarding the three outcomes of interest. To account
for potential measured confounding we developed separate large-scale
propensity score models within each database based on LASSO logistic
regression using the same predefined set of measured covariates
[@Tian2018]. To assume that treatment and comaparator cohorts stand in
equipoise we required that the majority of the patients in both have
preference scores between 0.3 and 0.7. We assumed that covariate
balance had been reached, if all after-adjustment baseline
characteristics had standardized absolute mean differences of less
than 0.1.  We estimated treatment effects using Cox proportional
hazards models with treatment as the sole covariate. To account for
observed confounding we used variable ratio matching (1:10) on the
propensity score using a caliper of 0.2 on the logit scale.  Hazard
ratios (HR) derived in different databases were summarized using
random effects meta-analysis.

For the second set of analyses, we used the standardized framework for
risk-based assessment of treatment effect heterogeneity, that follows
the directives of the PATH statement [@Kent2019; @KentEnE2019].
Following data extraction, the framework was applied in
two steps. First, we derived individualized risk predictions for the
three efficacy outcomes. We built the prediction models using LASSO
logistic regression on the propensity score matched (1:10)
subpopulation of the pooled treatment arms, aiming to remove any
effect of measured confounding that may cause differential fit of our
models across treatment arms. We considered the same large set of
candidate covariates as for the development of the propensity score
models.

For each outcome we used the derived prediction models to divide the
population in 4 equally-sized subgroups, based on the quarters of the
predicted risk distribution. Within each of these risk-based subgroups
we developed a new propensity score model. Our analyses were performed
on the propensity score matched (1:4) subset of the risk subgroup
subset. We derived relative effect estimates using Cox proportional
hazards models only with treatment as a predictor. Absolute effect
estimates were calculated based on the difference of the Kaplan-Meier
estimates, on day 365 after treatment initiation.

For each outcome we used the derived prediction models to stratify the
study population using two separate stratifcation schemes. In the
first approach, we divide the population into the lowest 75\% and the
top 25\% hip fracture risk subgroups. In the second approach, the
population is divided using existing guidelines on teriparatide
treatment based on major fracture risk (REF!!). Within each of these
risk-based subgroups we developed a new propensity score model. Our
analyses were performed on the propensity score matched (1:10) subset
of the risk subgroup subset. We derived relative effect estimates
using Cox proportional hazards models only with treatment as a
predictor. Absolute effect estimates were calculated based on the
difference of the Kaplan-Meier estimates, on day 365 after treatment
initiation.

Residual study bias from unmeasured confounding can still be present
in observational studies, which often may not be visible when
evaluating propensity score adjustment. To account for that, we
considered a set of 126 negative control outcome experiments, where a
null effect was assumed to be true. We used the estimated relative
effects for these outcomes to derive an empirical approximation to the
true null distribution, which was then used to calibrate the hazard
ratios and their 95% confidence intervals for the 3 outcomes of
interest [@Schuemie2013; @Schuemie2016].

\newpage
# Results
## Treatment effect heterogeneity

***[ Introductory stuff for Table 1 !! ]***

```{r, echo=FALSE, warning=FALSE, message=FALSE}
metaCalibrateOverall <- readRDS(here("data/processed/metaCalibrateOverall.rds"))

suffix <- "age_50_tr_1_q_25_75"

calibrateOverallResults <- readRDS(
    here(
      file.path(
          "data/processed",
          paste0(
              paste(
                  "calibrateOverallResults",
                  suffix,
                  sep = "_"
              ),
              ".rds"
          )
      )
    )
)

```

In all databases adequate equipoise of the preference score
distributions was achieved (Figure \ref{fig:overallPsDensity}). After
matching all baseline covariates (>35,000 in each database) were well
balanced (Figure \ref{fig:overallCovariateBalance}). This indicates
that we were able account for the majority of the observed
condfounding present in all databases. However, negative control
analyses time falsely favored oral bisphosphonates over teriparatide,
which suggests the presence of residual unobserved confoundng (Figure
\ref{fig:overallNcPlot}).

In terms of hip fracture risk we estimated calibrated hazard ratios of
`r calibrateOverallResults %>% filter(database == "ccae", outcome == 101) %>% pull(hr)`
(`r calibrateOverallResults %>% filter(database == "ccae", outcome == 101) %>% pull(lower)`
to
`r calibrateOverallResults %>% filter(database == "ccae", outcome == 101) %>% pull(upper)`;
95\% CI), 
`r calibrateOverallResults %>% filter(database == "optum_extended_dod", outcome == 101) %>% pull(hr)`
(`r calibrateOverallResults %>% filter(database == "optum_extended_dod", outcome == 101) %>% pull(lower)`
to
`r calibrateOverallResults %>% filter(database == "optum_extended_dod", outcome == 101) %>% pull(upper)`;
95\% CI) and
`r calibrateOverallResults %>% filter(database == "optum_ehr", outcome == 101) %>% pull(hr)`
(`r calibrateOverallResults %>% filter(database == "optum_ehr", outcome == 101) %>% pull(lower)`
to
`r calibrateOverallResults %>% filter(database == "optum_ehr", outcome == 101) %>% pull(upper)`;
95\% CI), in CCAE, Optum-DOD and Optum-EHR respectively. The
calibrated meta-analytic hazard ratio of the overall effect of
teriparatide compared to oral bisphosphonates was
`r metaCalibrateOverall %>% filter(outcome == 101) %>% pull(hr)`
(`r metaCalibrateOverall %>% filter(outcome == 101) %>% pull(lower)`
to
`r metaCalibrateOverall %>% filter(outcome == 101) %>% pull(upper)`;
95\% CI). Results for major osteoporotic and vertebral fractures are
provided in the supplement.

When using the internally developed hip fracture risk prediction models to divide the population into lower 75\% and top 25\% risk subgroups, we observed....

\newpage
# Discussion

\newpage
# References
\setlength{\parindent}{-0.25in}
\setlength{\leftskip}{0.25in}
\noindent
<div id="refs"></div>
\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}
\noindent

\newpage
# Tables and figures
\newpage
```{r overallPsDensity, cache = TRUE, echo = FALSE, warning=FALSE, fig.cap="Calibrated overall results"}
grid::grid.raster(tiff::readTIFF(here("figures/OverallPsDensity.tiff")))
```

\newpage
```{r overallCovariateBalance, cache = TRUE, echo = FALSE, warning=FALSE, fig.cap="Calibrated overall results"}
grid::grid.raster(tiff::readTIFF(here("figures/OverallCovariateBalance.tiff")))
```

\newpage
```{r overallNcPlot, cache = TRUE, echo = FALSE, warning=FALSE, fig.cap="Overall negative control plot."}
grid::grid.raster(tiff::readTIFF(here("figures/OverallNcPlot.tiff")))
```

\newpage
```{r plotMeta, cache=TRUE, echo=FALSE, warning=FALSE, fig.cap="Calibrated hazard ratios for the 3 outcomes of interest across the 3 considered databases using a set of 126 negative controls. Values below 1 favour teriparatide, while values above 1 favour bisphosphonates.", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here::here("figures/plotMeta.tiff")))
```

\newpage
```{r plotMetaRiskStratified, cache=TRUE, echo=FALSE, warning=FALSE, fig.cap="Calibrated hazard ratios for the 3 outcomes of interest across the 3 considered databases using a set of 126 negative controls. Values below 1 favour teriparatide, while values above 1 favour bisphosphonates.", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here::here("figures/plotMetaRiskStratified.tiff")))
```

\newpage
```{r plotAbsoluteRiskStratified, cache=FALSE, echo=FALSE, warning=FALSE, fig.cap="Calibrated hazard ratios for the 3 outcomes of interest across the 3 considered databases using a set of 126 negative controls. Values below 1 favour teriparatide, while values above 1 favour bisphosphonates.", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here::here("figures/plotAbsoluteHip.tiff")))
```